Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-05 vesicle_code/docs/adhesion/adhesion.tex
worked through shear example

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a shear flow}
\label{sec:sflow} 
We consider two vesicles suspended in the shear flow $\uu = \chi(y,0)$,
where $\chi$ is the shear rate.  The vesicles are placed on the $y$ axis
so that the background velocity drives the vesicles towards one another.
In the absence of adhesion, once the vesicles are sufficiently close,
they are deflected to opposite sides of the $y$ axis, and the result is
that the distance between the vesicles goes to infinity.  However, in
the presence of adhesion, is it possible that the adhesive force keeps
the vesicles from separating, resulting in the formation of a {\em
doublet}~\cite{}.

The formation of the doublet depends on four main parameters: the
vesicles' reduced area, the adhesion length scale $\delta$, the adhesion
strength $\mathcal{H}$, and the shear rate $\chi$.
Figure~\ref{fig:doublet090} shows the configuration of two vesicles that
have formed a doublet at several instances in time.  In
Figure~\ref{fig:sflow_phase_diagram} (left), we show the minimum
distance between the two vesicles as a function of time for various
adhesion strengths.  The adhesion strength corresponding to the red
curves is less than a critical strength $\mathcal{H}_c$ for a doublet to
form.  In contrast, for larger adhesive strengths, doublets are formed
(blue curves).

\begin{figure}[htp]
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image01.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image02.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image03.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image04.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image05.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image06.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image07.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image08.png}
  \caption{\label{fig:doublet090} The formation of a doublet in a shear
  flow.  The reduced area is $\Delta A = 0.90$, the Hamaker constant is
  $\mathcal{H}=0.7$, the adhesion length scale is $\delta = 0.4$, and
  the shear rate is $\chi=0.5$.}
\end{figure}

\begin{figure}
  \includegraphics[height=0.4\textwidth]{figs/shear_adR4em1Chi1e0_ra-090.pdf}
  \includegraphics[height=0.4\textwidth]{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \caption{{\em Left}: Distance between a pair of vesicles in a planar
  shear flow with shear rate $\chi=0.5$ and adhesion length scale
  $\delta = 0.4$.   The different lines correspond to a linear spacing
  of adhesion strengths ranging from $\mathcal{H}=0.1$ and
  $\mathcal{H}=1.0$.  {\em Right}: Phase diagram of vesicle
  hydrodynamics in a shear flow.  At a fixed reduced area $\Delta
  A=0.90$ and $\delta =0.4$,  the critical vdW potential strength
  $\mathcal{H}$ for binding of two vesicles in a shear flow depends on
  the shear rate $\chi$.
\label{fig:sflow_phase_diagram}}
\end{figure}

We investigate rheology of suspensions by computing the effective
viscosity of a doublet and compare it to the effective viscosity of a
single tank-treading vesicle. The effective viscosity is defined as the
viscosity of a homogeneous Newtonian fluid with the same energy
dissipation per macroscopic element of fluid.  In a simple shear flow,
the ambient velocity field is given by $\uu(\xx) = \chi(y,0)$, where
$\chi$ is the shear rate, and the intrinsic viscosity is
\begin{align*}
  [\mu]:= \frac{\mu_{\mathrm{eff}} - \mu_0}{\phi \mu_0} = 
  \frac{1}{\chi \mu_0 T} \int_{T_i}^{T_e} 
  \langle \sigma_{12} \rangle dt,
\end{align*}
where
\begin{align*}
  \langle \tau \rangle = \frac{1}{|\omega|} \int_{\gamma}
    \xxi \otimes \xx ds.
\end{align*}
Here, $[\mu]$ is the intrinsic viscosity, $\phi$ is the area fraction of
vesicles, $/tau$ is the stress due to the vesicles, and $\langle \cdot
\rangle$ is the spatial average.

In the left plot of Figure~\ref{fig:shearIntrinsicViscosity}, we compare
the intrinsic viscosity of a single tank treading vesicle to a doublet.
To validate our simulations, we superimpose (black marks) the intrinsic
viscosity calculated by Ghigliotti et
al.~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
significantly increases the intrinsic viscosity at all the reduced
areas.  To further characterize the effect of adhesion, in the right
plot of Figure~\ref{fig:shearIntrinsicViscosity} we decompose the
intrinsic viscosity into the contributions from the bending and tension
(blue), and the contribution from the adhesion (red).  We also
superimpose twice the intrinsic viscosity of a single tank treading
vesicle (dashed curve) to demonstrate that the bending and tension of
the doublet behaves similarly, but not identically, to a dilute
suspensions of non-adhering tank-treading vesicles.  We see that the
effect of the adhesion on the intrinsic viscosity is largest for
vesicles with small reduced areas.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} {\em Left}: The intrinsic
  viscosity of a tank treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\mathcal{H} = 0.7$,
  and the adhesion length scale is $\delta = 0.4$.  The length of the
  vesicles at each of the reduced areas is fixed, while the area changes
  with differing reduced areas.  The black marks denote intrinsic
  viscosity values computed by Ghigliotti et
  al.~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  {\em Right}:
  The decomposition of the intrinsic viscosity into the contributions
  for the bending and tension (blue) of the vesicles, and the
  contribution from the vesicle adhesion (red).  Also included is twice
  the intrinsic viscosity of a single tank-treading vesicle.}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a shear flow}
\label{sec:sflow} 
We consider two vesicles suspended in the shear flow $\uu = \chi(y,0)$,
where $\chi$ is the shear rate.  The vesicles are placed on the $y$ axis
so that the background velocity drives the vesicles towards one another.
In the absence of adhesion, once the vesicles are sufficiently close,
they are deflected to opposite sides of the $y$ axis, pass one another,
and separate from one another.  However, in the presence of adhesion,
the vesicles can form a doublet for certain values of the adhesion
length scale $\delta$, Hamaker constant $\mathcal{H}$, shear rate
$\chi$, and reduced area $\Delta A$.  Figure~\ref{fig:doublet090} shows
snapshots two vesicles that have formed a doublet and the color coding
is the tension along the vesicle.  The two vesicles are
identical with reduced area is $\Delta A = 0.9$, the shear rate is $\chi
= 0.5$, the adhesion length scale is $\delta = 0.4$, and the Hamaker
constant is $\mathcal{H} = 0.7$.  Similar to the quiescent example, we
observe that the membrane tension in the contact region is negative in
the contact region indicating that the membrane is being compressed when
the adhesive force is strongest.

\begin{figure}[htp]
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image01.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image02.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image03.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image04.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image05.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image06.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image07.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image08.png}
  \caption{\label{fig:doublet090} The formation of a doublet in a shear
  flow.  The reduced area is $\Delta A = 0.9$, the Hamaker constant is
  $\mathcal{H}=0.7$, the adhesion length scale is $\delta = 0.4$, and
  the shear rate is $\chi=0.5$.}
\end{figure}

We characterize the effect of the Hamaker constant and shear rate by
determining a critical Hamaker constant, $\mathcal{H}_c$, which
determines if the vesicles form a doublet or separate.  We fix the
adhesion length scale to $\delta = 0.4$ and the vesicle reduced area to
$\Delta A = 0.9$.  In the left plot of
Figure~\ref{fig:sflow_phase_diagram}, we plot the minimum distance
between the vesicles as a function of time for various Hamaker
constants.  The red curves correspond to Hamaker constants that are too
small to bind the vesicles into a doublet, and the blue curves
correspond to Hamaker constants that are sufficiently large to form a
doublet.

\begin{figure}
  \includegraphics[height=0.35\textwidth]{figs/shear_adR4em1Chi1e0_ra-090.pdf}
  \includegraphics[height=0.35\textwidth]{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \caption{{\em Left}: Distance between a pair of vesicles in a planar
  shear flow with shear rate $\chi=0.5$ and adhesion length scale
  $\delta = 0.4$.   The different lines correspond to a linear spacing
  of adhesion strengths ranging from $\mathcal{H}=0.1$ and
  $\mathcal{H}=1.0$.  {\em Right}: Phase diagram of vesicle
  hydrodynamics in a shear flow.  At a fixed reduced area $\Delta
  A=0.90$ and $\delta = 0.4$,  the critical Hamaker constant
  $\mathcal{H}$ for binding of two vesicles in a shear flow depends on
  the shear rate $\chi$.
\label{fig:sflow_phase_diagram}}
\end{figure}

Finally, we investigate the rheology of a suspension of a doublet by
computing the effective viscosity of a doublet and compare it to the
effective viscosity of a single tank-treading vesicle. The effective
viscosity is defined as the viscosity of a homogeneous Newtonian fluid
with the same energy dissipation per macroscopic element of fluid.  In a
simple shear flow, the intrinsic viscosity, $[\mu]$ is
\begin{align*}
  [\mu]:= \frac{\mu_{\mathrm{eff}} - \mu_0}{\phi \mu_0} = 
  \frac{1}{\chi \mu_0 T} \int_{T_i}^{T_e} 
  \langle \sigma_{12} \rangle dt,
\end{align*}
where
\begin{align*}
  \langle \tau \rangle = \frac{1}{|\omega|} \int_{\gamma}
    \xxi \otimes \xx ds,
\end{align*}
where $\phi$ is the area fraction of vesicles, $/tau$ is the stress due
to the vesicles, and $\langle \cdot \rangle$ is the spatial average.

In the left plot of Figure~\ref{fig:shearIntrinsicViscosity}, we compare
the intrinsic viscosity of a single tank treading vesicle to a doublet.
To validate our simulations, we superimpose (black marks) the intrinsic
viscosity calculated by Ghigliotti et
al.~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
significantly increases the intrinsic viscosity at all the reduced
areas.  To further characterize the effect of adhesion, in the right
plot of Figure~\ref{fig:shearIntrinsicViscosity} we decompose the
intrinsic viscosity into the contributions from the bending and tension
(blue), and the contribution from the adhesion (red).  We also
superimpose twice the intrinsic viscosity of a single tank treading
vesicle (dashed curve) to demonstrate that the bending and tension of
the doublet behave similarly, but not identically, to a dilute
suspensions of non-adhering tank-treading vesicles.  We see that the
effect of the adhesion on the intrinsic viscosity is largest for
vesicles with small reduced areas.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} {\em Left}: The intrinsic
  viscosity of a tank treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\mathcal{H} =
  0.7$, and the adhesion length scale is $\delta = 0.4$, and the length
  of the vesicles are held fixed.  The black marks denote intrinsic
  viscosity values computed by Ghigliotti {\em et
  al.}~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  {\em
  Right}: The decomposition of the intrinsic viscosity into the
  contributions for the bending and tension (blue) of the vesicles, and
  the contribution from the vesicle adhesion (red).  Also included is
  twice the intrinsic viscosity of a single tank-treading vesicle
  (dashed black line).}
\end{figure}
